---
title: "ex2 facebook bus"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ex 2 Facebook Bus
First I create a dataset based on the seats on the bus. The connection rule is: communication is really possible with adjacent seats only, but any type of adjacency works: side, front, back, diagonal, even across the aisle.

## Load the csv sheet
```{r}
library(tidyverse)
df = read_csv("~/Desktop/McGill/ORGB/2022-ona-assignments/ex2/facebook_bus.csv")
library(dplyr)
## preprocessing
## change all the node type to character
df <- df %>%
        mutate(node_1 = as.character(node_1))
df <- df %>%
        mutate(node_2 = as.character(node_2))
```
## Create a network based on the relationship
```{r}
library("ggraph")
library("tidygraph")
nodes <- data.frame(nodes = c(1,2,3,4,5,6,"A","B","C","D"))
edges = df
network <- tbl_graph(nodes=nodes, edges=edges, directed=FALSE)
```

```{r}
## plotting
ggraph(network, layout = "graphopt") + 
  geom_edge_link() + 
  geom_node_point()+geom_node_text(aes(label = nodes),repel = TRUE)
  theme_graph()

```
## Calculate various measures of centrality
```{r}
## centrality degree
network_centrality_degree = network %>%
  activate(nodes) %>%
  mutate(degree = centrality_degree()) %>%
  as_tibble()

## centrality betweeness
network_betweenness_centrality = network %>%
  activate(nodes) %>%
  mutate(betw = centrality_betweenness()) %>%
  arrange(desc(betw)) %>% as_tibble()
```
```{r}
network_centrality_degree
```
```{r}
network_betweenness_centrality
```
## Choice of Seats
The \textbf{degree centrality} of a node is simply its degree—the number of edges it has.

\textbf{Betweenness centrality} measures the extent to which a vertex lies on paths between other vertices. Vertices with high betweenness may have considerable influence within a network by virtue of their control over information passing between others. 

In order to maximize the change of building a relationship between co-workers, I will need to pick a seat that can talk to as much people as possible, meaning a seats with high centrality degree. Also, I would like a seat with high betweenness centrality, because that seats will help me passing information to others.

If we calculate the "value" of a seats by adding the degree centrality and betweenness centrality. The seat that has the highest value will be B. This choice is beneficial if seats D,C and A will be filled. It will be even more beneficial if people who will seat at D,C or C are permanent worker at facebook. It is not so beneficial if people who seat at D,C or A are interns like me. It is even less benificial if no one will seats at D,C and A.

## Plot the network graph with labels and centrality values
```{r}
name <- c("1","2","3","4","5","6","A","B","C","D")
deg_cen <- c(1,2,5,2,3,3,3,5,5,5)
bwt_cen <- c(0,8,4.63,0,0.5333333,0.9333333,14,9.0333333,8.6000000,3.2666667)
nodes <- data.frame(name, deg_cen,bwt_cen)

network_2 <- tbl_graph(nodes=nodes, edges=edges, directed=FALSE)
```
### Plot based on degree centrality
```{r}
ggraph(network, layout = "graphopt") + 
  geom_edge_link() + 
  geom_node_point(aes(size = deg_cen))+geom_node_text(aes(label = nodes),repel = TRUE)
  theme_graph()
```
### Plotting based on betweenness centrality
```{r}
ggraph(network, layout = "graphopt") + 
  geom_edge_link() + 
  geom_node_point(aes(size = bwt_cen))+geom_node_text(aes(label = nodes),repel = TRUE)
  theme_graph()
```

