---
title: "ex3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data
```{r,warnings = FALSE}
library(arrow)
df = read_parquet("~/Desktop/McGill/ORGB/2022-ona-assignments/ex3/app_data_sample.parquet")
```
## Predicting examiners' gender based on first name:
The gender package attempts to infer gender (or more precisely, sex assigned at birth) based on first names using historical data, typically data that was gathered by the state.

```{r, warnings = FALSE}
library(gender)
library(dplyr)
first_name = df %>%distinct(examiner_name_first)
gender_probability = gender(first_name$examiner_name_first)
gender_dictionary = gender_probability %>% select(name,gender)
df <- df %>% left_join(gender_dictionary, by = c("examiner_name_first" = "name"))
head(df$gender)

```
The gender package assign gender based on historical data. Some of the name is not in the data set, thus there are some missing gender information. I filled those values by distribution.
```{r, warnings = FALSE}
table(is.na(df$gender))
gender_na = is.na(df$gender)
gender_fill = sample(df$gender[!gender_na], size = sum(gender_na), replace = TRUE)
df$gender[is.na(df$gender)] <- gender_fill
table(is.na(df$gender))
```
All the missing value has been filled.

## Predicting examiners' race based on last name:
The "predictrace" package predict the race of a surname using U.S. Census data which says how many people of each race has a certain surname.
```{r,warnings = FALSE}
library(predictrace)
race = predict_race(df$examiner_name_last, probability = FALSE)
df$race = race$likely_race
head(df$race,10)
```
Again, fill the missing values based on distribution.
```{r, warnings = FALSE}
table(is.na(df$race))
race_na = is.na(df$race)
race_fill = sample(df$race[!race_na], size = sum(race_na), replace = TRUE)
df$race[is.na(df$race)] <- race_fill
table(is.na(df$race))
```
## Calculate Tenure
To calculate tenure, I need to calculate the time the application stay in the system.

For most applications, the filing date is the date on which PTO received the application.

The appl_status_date variable indicates the date that the application entered its most recent status (or status as of the end of 2014). 
```{r, warnings = FALSE}
tenure_info <- df %>% select(examiner_id, filing_date, appl_status_date)

library(lubridate)
tenure_info = tenure_info %>% mutate(appl_status_date = as_date(dmy_hms(appl_status_date))) 
tenure_info$tenure_days = as.numeric(difftime(tenure_info$appl_status_date,tenure_info$filing_date,units="days"))

## detect missing values
table(is.na(tenure_info$tenure_days))

## fill missing values
tenure_na = is.na(tenure_info$tenure_days)
tenure_fill = sample(tenure_info$tenure_days[!tenure_na], size = sum(tenure_na), replace = TRUE)
tenure_info$tenure_days[is.na(tenure_info$tenure_days)] <- tenure_fill
table(is.na(tenure_info$tenure_days))

## join with df
df$tenure = tenure_info$tenure_days
```

## Pick two workgroup
The two group I pick is 1648 and 1722.
1600 – Biotechnology
1700 – Chemical and Materials Engineering
```{r}
wg = as.numeric(substr(df$examiner_art_unit, 1, 3))
df$wg = wg
group_164 = df %>% filter(df$wg == 164)
group_172 = df %>% filter(df$wg == 172)

```
### Examing Group 1648

```{r}
## summary
table(group_164$gender)
```

```{r}
table(group_164$race)
```


### Examing Group 1722

```{r}
## summary
table(group_172$gender)
```

```{r}
table(group_172$race)
```

```{r}
two_group_gender <- t(cbind(table(group_164$gender), table(group_172$gender)))

barplot(two_group_gender, beside=T, col=c("red","blue")) 
par(xpd=T)
legend("top",legend = c("164","172"), fill=c("red","blue"), cex = 0.5)
```
```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE,results='hide',fig.keep='all'}

two_group_race <- t(cbind(table(group_164$race), table(group_172$race)))

two_group_race[1,1] = 0

barplot(two_group_race, beside=T, col=c("red","blue")) 
par(xpd=T)
legend("top",legend = c("164","172"), fill=c("red","blue"), cex = 0.5)
```
## Create advice networks from edges-sample

```{r, message=FALSE}
library(tidyverse)
net = read_csv("~/Desktop/McGill/ORGB/2022-ona-assignments/ex3/edges_sample.csv")

edges_164 = inner_join(df %>% filter(wg == 164),net,by = c("application_number" = "application_number")) %>% select(ego_examiner_id,alter_examiner_id,examiner_art_unit)

colnames(edges_164) = c("from","to","art_unit")
edges_164 = drop_na(edges_164)

edges_172 = inner_join(df %>% filter(wg == 172),net,by = c("application_number" = "application_number")) %>% select(ego_examiner_id,alter_examiner_id,examiner_art_unit)

colnames(edges_172) = c("from","to","art_unit")
edges_172 = drop_na(edges_172)
```
### Create Nodes
```{r,message=FALSE}
edges = rbind(edges_164,edges_172)
node_ego = edges %>% select(from,art_unit) %>%rename(id=from) 
node_alter = edges %>% select(to,art_unit)%>%rename(id=to)
nodes_all <-rbind.data.frame(node_ego, node_alter)

nodes = nodes_all %>% distinct(id)

nodes = nodes %>% mutate(id = as.character(id))
```


### Create Graph
```{r,warning=FALSE}
library(igraph)

net_164 = graph_from_data_frame(d=edges_164, vertices=nodes, directed=TRUE)
net_164
net_172 = graph_from_data_frame(d=edges_172, vertices=nodes, directed=TRUE)
```

## Pick the mesure of centrality
1. Degree centrality is defined as the number of links incident upon a node
2. Eigenvector Centrality is an algorithm that measures the transitive influence of nodes. A high eigenvector score means that a node is connected to many nodes who themselves have high scores.
3. Closeness centrality is a measure of the average shortest distance from each vertex to each other vertex
4. Betweenness centrality is a way of detecting the amount of influence a node has over the flow of information in a graph.
```{r}
## Degree Centrality
V(net_164)$dc <- degree(net_164)
V(net_172)$dc <- degree(net_172)

## Eigenvector Centrality
V(net_164)$ec <- evcent(net_164)$vector
V(net_172)$ec <- evcent(net_172)$vector

## Closeness Centrality
V(net_164)$cc <- closeness(net_164)
V(net_172)$cc <- closeness(net_172)

## Betweenness Centrality
V(net_164)$bc <- betweenness(net_164)
V(net_172)$bc <- betweenness(net_172)

```
## Plot the network based on centrality
```{r}
library(ggraph)
library(ggplot2)
library(ggpubr)
# Degree Centrality
dc_164 = ggraph(net_164, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=dc), show.legend=T) + ggtitle("Degree Centrality 164")

# Eigenvector Centrality
ec_164<-ggraph(net_164, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=ec), show.legend=T) + ggtitle("Eigenvector Centrality 164")

# Closeness Centrality
cc_164<-ggraph(net_164, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=cc), show.legend=T) + ggtitle("Closeness Centrality 164")

# Betweenness Centrality
bc_164<-ggraph(net_164, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=bc), show.legend=T) + ggtitle("Betwenness Centrality 164")
```

## Centrality Scores
```{r}
centrality_164 <- data.frame(
                             id = V(net_164)$name,
                         degree      = V(net_164)$dc,
                         closeness   = V(net_164)$cc,
                         betweenness = V(net_164)$bc,
                         eigenvector = V(net_164)$ec)
centrality_164
```
```{r}
centrality_172 <- data.frame(id = V(net_164)$name,
                         degree      = V(net_172)$dc,
                         closeness   = V(net_172)$cc,
                         betweenness = V(net_172)$bc,
                         eigenvector = V(net_172)$ec)
centrality_172
```

```{r}
ggarrange(dc_164,ec_164,cc_164,bc_164,ncol = 2, nrow = 2)
```

```{r}
dc_172 = ggraph(net_172, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=dc), show.legend=T) + ggtitle("Degree Centrality 172")

# Eigenvector Centrality
ec_172<-ggraph(net_172, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=ec), show.legend=T) + ggtitle("Eigenvector Centrality 172")

# Closness Centrality
cc_172<-ggraph(net_172, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=cc), show.legend=T) + ggtitle("Closeness Centrality 172")

# Betwenness Centrality
bc_172<-ggraph(net_172, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=bc), show.legend=T) + ggtitle("Betweenness Centrality 172")
```


```{r}
ggarrange(dc_172,ec_172,cc_172,bc_172,ncol = 2, nrow = 2)
```


Based on the graph, seems like closeness centrality has clearer cluster center.

## Characterize and discuss the relationship between centrality and other examiners’characteristics

```{r}
ggraph(net_164, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=cc), show.legend=T) +geom_node_text(aes(label = centrality_164$id), repel=TRUE, size=2)+ ggtitle("Closeness Centrality 164")
```

```{r}
ggraph(net_172, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=cc), show.legend=T) +geom_node_text(aes(label = centrality_172$id), repel=TRUE, size=2)+ ggtitle("Closeness Centrality 172")

```
### Gather all examiner characteristics
```{r}
examiner = df %>% select(examiner_id,examiner_art_unit,gender,race,tenure)
examiner = distinct(examiner)
```


### Examiner that are in group 164 and has the highest closeness centrality
```{r,warning=FALSE}
max_cc_164 = max(centrality_164$closeness[!is.na(centrality_164$closeness)])
max_cc_164_id = centrality_164 %>% filter(centrality_164$closeness ==max_cc_164) %>%select(id)
max_cc_164_id = max_cc_164_id %>% mutate(id = as.numeric(id))
max_cc_164_info = examiner %>%filter(examiner_id == max_cc_164_id$id)

table(max_cc_164_info$gender)

table(max_cc_164_info$race)
```
Examiners that has higher closeness centrality in group 164, are more likely to be while male.

### Examiner that are in group 172 and has the highest closeness centrality
```{r,warning=FALSE}
max_cc_172 = max(centrality_172$closeness[!is.na(centrality_172$closeness)])
max_cc_172_id = centrality_172 %>% filter(centrality_172$closeness ==max_cc_172) %>%select(id)
max_cc_172_id = max_cc_172_id %>% mutate(id = as.numeric(id))
max_cc_172_info = examiner %>%filter(examiner_id == max_cc_172_id$id)

table(max_cc_172_info$gender)

table(max_cc_172_info$race)
```
The examiners that has higher closeness centrality in group 172 are mostly male comparing to group 164. Also, there are more Hispanic examiners that are influencial in this group.