---
title: "ex 1: linkedin network"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the File

Download linkedin network file and import to R:

```{r}
library(tidyverse)
library(janitor)
df = read_csv("/Users/hong/Desktop/McGill/ORGB/Connections.csv")
df <- clean_names(df)
df
```

## Data Cleaning

Some company start with mcgill but has specific department, I will change all of them to just mcgill university". Similar for Ubisoft. Some people label their company as Ubisoft while others add specific studios after ubisoft.
```{r}
df <- df %>% 
  mutate(company = replace(company, str_detect(company, "McGill University"), "McGill University"))

df <- df %>% 
  mutate(company = replace(company, str_detect(company, "Ubisoft"), "Ubisoft"))
```
## Total number of contact:
```{r}
length(df$first_name)
```


## Plot the Counting result
```{r}
library(dplyr)
company_count = df %>% count(company, sort = TRUE) ## count the number of ppl in each company

## plot the counts
company_count %>% ggplot(aes(y = reorder(company,n), x=n))+
  geom_col(fill="grey") +theme(text = element_text(size = 8)) +
  labs(
    x = "Connection Counts",
    y = "Company"
  )
```
## Create Nodes:

```{r}
## Concat first name last name
name =  paste0(df$first_name,"_", df$last_name)
df$name = name
nodes <- df %>% select(name)
```
## Adding Connection Edges:
```{r}
## Concat first name last name

df1 <- data.frame(df$name,df$company)
df2 <- data.frame(df$name,df$company)
edges <- full_join(df1, df2, by = "df.company")
names(edges) <- c("from", "company", "to")


```
## Linkedin Network
```{r}
library("ggraph")
library("tidygraph")
network <- tbl_graph(nodes=nodes, edges=edges, directed=FALSE)
network

```
## Plot:
```{r}
ggraph(network, layout = "graphopt") + 
  geom_edge_link(aes(color = company), show.legend = FALSE) + 
  geom_node_point()+
  theme_graph()
```


